---
title: "Exploratory analysis for memorability of food stimuli"
author: "Xinyue Li"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/Christine/Desktop/python/mem_dm")
library(MASS)
library(plyr)
library(corrplot)
library(RColorBrewer)
suppressWarnings(library(ggplot2))
library(tidyverse)
library(nFactors)
dat.f = read.csv("food_analysis/stimchar_final_touse/stimchar_final_touse.csv")
dat.f = dat.f %>% filter(age == "18-25" | age == "26-30" | age == "31-35")
name = read.csv("mem.csv")

```

## Set up
The food folios data is filtered by age (18-35).
Same as what mentioned in the paper, we will also flip 4 of the measures:

Sugar-low sugar

Disgusting-not disgusting

Healthy-unhealthy

Vitamins-low vitamins

(All of the following codes are based on source code by Shehzad on github: [github.com/czarrar/food_choice](https://github.com/czarrar/food_choice/blob/main/20_factor_scores.Rmd))
 
```{r rdat}
rdat = dat.f %>% select(stimulus, choice.rating, starts_with("rating"))
colnames(rdat) = sub("rating[.]", "", colnames(rdat))
colnames(rdat) = sub("[.]rating", "", colnames(rdat))

rdat$sugar = 10 - rdat$sugar
rdat$disgusting = 10 - rdat$disgusting
rdat$healthy = 10 - rdat$healthy
rdat$vitamins = 10 - rdat$vitamins

# rename
cinds = sapply(c("sugar", "disgusting", "healthy", "vitamins"), function(x) which(colnames(rdat) == x))
colnames(rdat)[cinds] = c("low-sugar", "not-disgusting", "unhealthy", "low-vitamins")

rmat = as.matrix(rdat[,-1])
rownames(rmat) = rdat$stimulus
rmat.nochoice = rmat[,-1]
```

## Correlation Matrix
Correlation Matrix of filtered original data and memorability score
The patterns of ratings are roughly the same as the correlation matrix in the manuscript.
Memorability showed positive correlation with **tasty**, **other tasty**, **feel**, which are major loadings for the factor **tasty** analyzed in the manuscript.
```{r ave.mat}
# average of ratings for each image
# follow the order of values on paper
new.ord = c("tasty", "othertasty", "feel", "texture", "not-disgusting", "familiar", "filling", "unhealthy", "sweetsaltysavory", "calories", "carbohydrates", "fat", "low-vitamins", "gluten", "sodium", "low-sugar", "protein", "stimulus")

# correlation matrix with memorability score
ave.mat.r = rdat[,new.ord] %>% 
  group_by(stimulus) %>% 
  summarise_all(mean) 
# matching memorability stimulus name to food data stimulus name
ave.mat.r = data.frame(Memorability = name$Memorability, mem.stimulus = name$X, ave.mat.r)

M.mem = ave.mat.r %>% select(-mem.stimulus, -stimulus) %>% cor()
diag(M.mem) = 0
corrplot(M.mem, method="circle", diag = F, #order="hclust", 
         #is.corr=F, cl.lim=c(-0.85,0.85), 
         col=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(256)))
```

# Factor Analysis

Run factor analysis with the filtered food data. The results are the same as the original dataset. There are 3 factors,and the loads of factors are showing same pattern as the paper showed.

Three factors are named as: **Tasty, Healthiness, and Savory**.

Plot loading of ratings to the three factors with the filtered data.
```{r }
# factor analysis
nFactors::nCng(as.data.frame(rmat.nochoice), model="factors") 
fit <- factanal(rmat.nochoice, 3, scores="regression")

d <- dist(fit$loadings)
hc <- hclust(d, method="average")
dend <- as.dendrogram(hc)
loads <- fit$loadings[order.dendrogram(dend),]
loads <- rbind(choice=loads[rownames(loads) == "choice",], loads[rownames(loads) != "choice",])
# Make tasty first, then health, and finally sweet/savory
loads2 <- loads[,c(2,1,3)]
colnames(loads2) <- sprintf("Factor-%i", 1:3)
# Put the sweet/savory to the end
loads2 <- rbind(loads2[-c(1:4),], loads2[1:4,])
colnames(loads2) = c("Tasty", "Healthiness", "Savory")
# Plot
 cols2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", 
                            "#F4A582", "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE", "#4393C3", "#2166AC", "#053061"))(200)
corrplot::corrplot(loads2, diag=T, col=rev(cols2), cl.pos="n")

```

## Multiple Regression
The multiple regression result showed that **tasty** is the only factor that can predict memorability.

Show as the plot below:
```{r warning=FALSE}
scores3 <- fit$scores
ind1 <- which.max(fit$loadings[rownames(fit$loadings) == "unhealthy",])
ind2 <- which.max(fit$loadings[rownames(fit$loadings) == "tasty",])
ind3 <- which.max(fit$loadings[rownames(fit$loadings) == "sweetsaltysavory",])
colnames(scores3) <- c("Food.UnHealth", "Food.Taste", "Food.SweetProtein")[c(ind1,ind2,ind3)]
outdf <- data.frame(subjectId=dat.f$subjectId, stimulus=dat.f$stimulus, scores3)

# score of factor for each food item
ave.factor = outdf %>% group_by(stimulus) %>% 
  summarise_all(mean) 

# add mem score to the factor data frame
labels = sapply(name$stimulus, function(x) which(ave.factor$stimulus == x))
ave.factor.mem = data.frame(Memorability = name$Memorability, mem.stimulus = name$X, ave.factor)

# perform linear regression
fit1 = lm(Memorability ~ Food.Taste, data = ave.factor.mem)
summary(fit1)

# plot mem~food.taste
ave.factor.mem %>% ggplot(aes(x = Food.Taste, y = Memorability))+
  geom_point()+
  geom_smooth(method='lm', se = FALSE)
```





